---
title: "Module 4 - Instructions"
author: "Oliver Engist"
date: "04/04/2021"
output: pdf_document
---

In the next assignment we want to replicate some plots from the paper "Female Socialization: How Daughters Affect Their Legislator Fathers' Voting on Women's Issues" (Washington, 2008). The paper explores whether having a daughter makes politicians more sensitive to women's rights issues and how this is reflected in their voting behavior. The main identifying assumption is that after controlling for the number of children, the gender composition is random. This might be violated if families that have a preference for girls keep having children until they have a girl. In this assignment we will prepare a dataset that allows us to test whether families engage in such a "female child stopping rule".


# Setup

* Load the libraries "Rio" and "tidyverse"
* Change the path of the working directory to your working directory.

```{r Load Libraries, message=FALSE}
library(rio)
library(tidyverse)
```

* import the data sets *basic.dta* and *genold108.dta*
```{r Import_Data, message=FALSE}
basic_data <- import("basic.dta")
genold_108 <- import("genold108.dta")
```

* create a subset of the 108th congress from the *basic* dataset
```{r Subset Creation, message=FALSE}
sub_108_data <- subset(basic_data, congress=="108")
```

* join this subset with the *genold* dataset
```{r Join Data, message=FALSE}
genold_data <- left_join(x=sub_108_data, y=genold_108, by=c("name","statenam"))
```

# Data preparation

* check table 1 in the appendix of the paper and decide which variables are necessary for the analysis (check the footnote for control variables)
# Control Variables are legislator race, gender, party, age, age squared, service length, service length square, religion, region, total number of children. Additionally include number of daughters, Party.

* drop all other variables.
```{r Slected relevant variable, message=FALSE}
appendix_var <- select(genold_data, white, female, party, age, srvlng, rgroup, region, totchi, ngirls, anygirls, demvote, genold)
```

* Recode *genold* such that gender is a factor variable and missing values are coded as NAs.
#while col was a character, I changed all blanks to NA. Then set to factor.
```{r genold factors, message=FALSE}
appendix_var$genold[appendix_var$genold==""] <- NA
appendix_var$genold <- factor(appendix_var$genold)
```

* Recode *party* as a factor with 3 levels (D, R, I)
I think var 2 is republican - checked against another column. I've assigned I to democrats as we're asked to do this at the end.
```{r party with 3 variable, message=FALSE}
appendix_var$party <- factor(appendix_var$party, levels = c(1,2,3), labels = c("D","R","D"))
```

* Recode *rgroup* and *region* as factors.
```{r recode factors}
appendix_var$rgoup <- factor(appendix_var$rgroup)
appendix_var$region <- factor(appendix_var$region)
```

* generate variables for age squared and service length squared
```{r - add variables with mutuate function}
appendix_var <- appendix_var %>% mutate(age_sqrd=age^2)
appendix_var <- appendix_var %>% mutate(srvlng_sqrd=srvlng^2)
```

* create an additional variable of the number of children as factor variable

```{r - number of children as factor varialbe}
appendix_var <- appendix_var %>% mutate(numchildren_fct=as_factor(totchi))
```

# Replicationg Table 1 from the Appendix

We haven't covered regressions in R yet. Use the function *lm()*. The function takes the regression model (formula) and the data as an input. The model is written as $y \sim x$, where $x$ stands for any linear combination of regressors (e.g. $y \sim x_1 + x_2 + female$). Use the help file to understand the function.

* Run the regression $total.children = \beta_0 + \beta_1 gender.oldest + \gamma'X$ where $\gamma$ stands for a vector of coefficients and $X$ is a matrix that contains all columns that are control variables.\footnote{This is just a short notation instead of writing the full model with all control variables $totchi = \beta_0 + \beta_1 genold + \gamma_1 age + \gamma_2 age^2 + \gamma_3 Democrat + ... + \epsilon$ which quickly gets out of hand for large models.}
```{r Run regression for total children}
Regression_totchi <- lm(totchi ~ white + female + party + age + age_sqrd + srvlng + srvlng_sqrd + rgroup + region + genold, data = appendix_var)
```

* Save the main coefficient of interest ($\beta_1$)
```{r save coefficient}
Beta_genold_totchi <- coef(summary(Regression_totchi))["genoldG","Estimate"]
Std_genold_totchi <- coef(summary(Regression_totchi))["genoldG","Std. Error"]
```

* Run the same regression separately for Democrats and Republicans (assign the independent to one of the parties). Save the coefficient and standard error of *genold*

```{r Running regression of democrat, republican and standard errors}
Regression_dem <- lm(totchi ~ white + female + age + age_sqrd + srvlng + srvlng_sqrd + rgroup + region + genold, data = appendix_var, party=="D")

Regression_republican <- lm(totchi ~ white + female + age + age_sqrd + srvlng + srvlng_sqrd + rgroup + region + genold, data = appendix_var, party=="R")

Beta_genold_dem <- coef(summary(Regression_dem))["genoldG","Estimate"]
Beta_genold_rep <- coef(summary(Regression_republican))["genoldG","Estimate"]

Std_genold_dem <- coef(summary(Regression_dem))["genoldG","Std. Error"]
Std_genold_rep <- coef(summary(Regression_republican))["genoldG","Std. Error"]
```

* Collect all the *genold* coefficients from the six regressions, including their standard errors and arrange them in a table as in the paper.
```{r collect number of children for the N row and then repeat above steps for the number of daughters regression}
n.children_tot <- nobs(Regression_totchi)
n.children_dem <- nobs(Regression_dem)
n.children_rep <- nobs(Regression_republican)

Regression_ngirls <- lm(ngirls ~ white + female + party + age + age_sqrd + srvlng + srvlng_sqrd + rgroup + region + genold, data = appendix_var)

Regression_dem_girls <- lm(ngirls ~ white + female + age + age_sqrd + srvlng + srvlng_sqrd + rgroup + region + genold, data = appendix_var, party=="D")

Regression_republican_girls <- lm(ngirls ~ white + female + age + age_sqrd + srvlng + srvlng_sqrd + rgroup + region + genold, data = appendix_var, party=="R")

Beta_genold_dem_girls <- coef(summary(Regression_dem_girls))["genoldG","Estimate"]

Beta_genold_ngirls <- coef(summary(Regression_ngirls))["genoldG","Estimate"]

Beta_genold_rep_girls <- coef(summary(Regression_republican_girls))["genoldG","Estimate"]

Std_genold_dem_girls <- coef(summary(Regression_dem_girls))["genoldG","Std. Error"]

Std_genold_rep_girls <- coef(summary(Regression_republican_girls))["genoldG","Std. Error"]

Std_genold_ngirls <- coef(summary(Regression_ngirls))["genoldG","Std. Error"]

n.daughter_ngirls <- nobs(Regression_ngirls)
n.daughter_dem <- nobs(Regression_dem_girls)
n.daughter_rep <- nobs(Regression_republican_girls)
```

```{r make table}
Appendix_1_table <- matrix(c(Beta_genold_ngirls,Beta_genold_totchi,Beta_genold_dem_girls,Beta_genold_dem,Beta_genold_rep_girls,Beta_genold_rep,Std_genold_ngirls,Std_genold_totchi,Std_genold_dem_girls,Std_genold_dem,Std_genold_rep_girls,Std_genold_rep,n.daughter_ngirls,n.children_tot,n.daughter_dem,n.children_dem,n.daughter_rep,n.children_rep),ncol=6, byrow = TRUE)

row.names(Appendix_1_table) <- c("First Child","Female", "Number")
colnames(Appendix_1_table) <- c("Full Congress - Daughters", "Full Congress - Total Children", "Democrats - Daughters", "Democrats - Total Children", "Republicans - Daughters", "Republicans - Total Children")

view(Appendix_1_table)
print(Appendix_1_table)
```

* print the table