---
title: "R - PS4"
author: "Lingzhi Huang"
date: "4/22/2020"
output:
  pdf_document: default
  html_document: default
---

# Setup
```{r, message=FALSE}

library(rio)
library(tidyverse)
library(knitr)
```
```{r}
basic<- import('basic.dta')
genold<- import('genold108.dta')
```

```{r}
basic108th <- subset(basic, basic$congress==108,)
```

```{r, message=FALSE}
Combinedtable <- inner_join(genold,basic108th, by='name')
print(Combinedtable)
```
```{r}
Combinedtable <- Combinedtable[,c('name','genold','party','ngirls','nboys','totchi','rgroup','region','repub','srvlng','female','white','age')]
print(Combinedtable)
```

# Data preparation

```{r}
Combinedtable$genold <- as.factor(Combinedtable$genold)
Combinedtable$genold[Combinedtable$genold == ''] <- NA
print(Combinedtable)
```

```{r}
Combinedtable$party <- factor(Combinedtable$party, levels = c(1,2,3), labels=c("D", "R", "I"))
Combinedtable$rgroup <- factor(Combinedtable$rgroup)
Combinedtable$region <- factor(Combinedtable$region)
print(Combinedtable)
```

generate variables for age squared and service length squared
```{r}
Combinedtable<- Combinedtable%>%mutate(Agesqr=age^2)
Combinedtable <- Combinedtable%>%mutate(Srvlngsqr=srvlng^2)
```
create an additional variable of the number of children as factor variable
```{r}
Combinedtable$childrenfactor<-as.factor(Combinedtable$totchi)
```

# Replicating Table 1 from the Appendix

regression on number of children
```{r}
lm1 <- lm(totchi~ genold + white + party +female+age + Agesqr + srvlng+ Srvlngsqr+ rgroup+ region, data = Combinedtable)
```

Save the main coefficient of interest
```{r}
coeffcient1<-coefficients(lm1)
coeffcient1<-as.matrix(coeffcient1)
beta1<-coeffcient1[2,1]
summary1<-summary(lm1)
sucoe1<-summary1$coefficients[2,2]
```

regression on number of children
```{r}
lm2<-lm(ngirls~ genold+ party+ rgroup + region
         + srvlng + female + white + age + Agesqr
         + Srvlngsqr + childrenfactor, data=Combinedtable)

summary2<-summary(lm2)
beta12<-summary2$coefficients[2,1]
se_2<-summary2$coefficients[2,2]

```

Run the same regression separately for Democrats and Republicans (assign the independent to one of the parties). Save the coefficient and standard error of *genold*

```{r}
republicans<-Combinedtable[Combinedtable$repub==1,]
democrats<-Combinedtable[Combinedtable$repub==0,]

lm3<-lm(totchi~ genold+ rgroup + region
           + srvlng + female + white + age + Agesqr 
           + Srvlngsqr , data=republicans)

lm4<-lm(totchi~ genold+ + rgroup + region 
           + srvlng + female + white + age + Agesqr
           + Srvlngsqr , data=democrats)

 
summary4<-summary(lm4)
beta14<-summary4$coefficients[2,1]
se_4<-summary4$coefficients[2,2]

summary3<-summary(lm3)
beta13<-summary3$coefficients[2,1]
se_3<-summary3$coefficients[2,2]
```

```{r}
lm5<-lm(ngirls~ genold+ + rgroup + region
            + srvlng + female + white + age + Agesqr
            + Srvlngsqr + childrenfactor , data=democrats)

lm6<-lm(ngirls~ genold+ rgroup + region 
            + srvlng + female + white + age + Agesqr 
            + Srvlngsqr + childrenfactor , data=republicans)


summary5<-summary(lm5)
beta15<-summary5$coefficients[2,1]
se_5<-summary5$coefficients[2,2]

summary6<-summary(lm6)
beta16<-summary6$coefficients[2,1]
se_6<-summary6$coefficients[2,2]

table<-matrix(c(beta1,beta12,beta13,
                beta14,beta15,beta16,se,se_2,se_3,
                se_4,se_5,se_6),ncol=6,nrow=2,byrow=TRUE)

rownames(table)<-c("First child female","s.e.")
colnames(table)<-c("N girls full congress", "N children full congress", "N girls dem.", "N children dem.","N girls rep.","N children rep.")
```

* print the table
```{r}
print(table)
```
















